---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "READMEfigs_",
  cache = TRUE
)
```

# Mcomp: Data from the M-Competitions

[![Travis-CI Build Status](https://travis-ci.org/robjhyndman/Mcomp.svg?branch=master)](https://travis-ci.org/robjhyndman/Mcomp)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/Mcomp)](https://cran.r-project.org/package=Mcomp)
[![Downloads](http://cranlogs.r-pkg.org/badges/Mcomp)](https://cran.r-project.org/package=Mcomp)
[![Licence](https://img.shields.io/badge/licence-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)

The R package *Mcomp* contains the 1001 time series from the M-competition (Makridakis et al. 1982) and the 3003 time series from the IJF-M3 competition (Makridakis and Hibon, 2000).

## Installation

You can install the **stable** version on
[R CRAN](https://cran.r-project.org/package=Mcomp).

```r
install.packages('Mcomp')
```

You can install the **development** version from
[Github](https://github.com/robjhyndman/Mcomp)

```r
# install.packages("devtools")
devtools::install_github("robjhyndman/Mcomp")
```

## Usage

```{r usage}
library(Mcomp)
library(ggplot2)
M1
autoplot(M1$YAF2)
subset(M1,"monthly")
```

## Sources

[Makridakis, S., A. Andersen, R. Carbone, R. Fildes, M. Hibon, R. Lewandowski, J. Newton, E. Parzen, and R. Winkler (1982) The accuracy of extrapolation (time series) methods: results of a forecasting competition. *Journal of Forecasting*, **1**, 111--153.](http://doi.org/10.1002/for.3980010202)

[Makridakis and Hibon (2000) The M3-competition: results, conclusions and implications. *International Journal of Forecasting*, **16**, 451-476.](https://doi.org/10.1016/S0169-2070(00)00057-1)

## License

This package is free and open source software, licensed under GPL-3
